---
title: "HW5"
author: "NZ"
date: "2025-03-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Homewrok5

Nicolás Zapata

### 1
Assign to the variable n_dims a single random integer between 3 and 10.

```{r, eval=TRUE, echo=TRUE}
n_dims <- sample(3:10, size=1)
print(n_dims)

```

Create a vector of consecutive integers from 1 to n_dims2
```{r, eval=TRUE, echo=TRUE}
my_vector <- seq(1:n_dims^2)

```

Use the sample function to randomly reshuffle these values.
```{r, eval=TRUE, echo=TRUE}
matrix1 <- matrix(sample(my_vector), nrow=n_dims)
# print out the matrix
print(matrix1) 
```

Find a function in r to transpose the matrix
``` {r, eval=TRUE, echo=TRUE}
transposed_matrix1 <- t(matrix1)
print(transposed_matrix1)
```

Calculate the sum and the mean of the elements in the first row and then the last row.
```{r, eval=TRUE, echo=TRUE}
# mean first row
firstrow_mean <- mean(matrix1[1,])
print(firstrow_mean) 
# sum  first row
firstrow_sum <- sum(matrix1[1,])
print(firstrow_sum)
```

```{r, eval=TRUE, echo=TRUE}
# mean last row 
lastrow_mean <- mean(matrix1[n_dims,])
print(lastrow_mean)
# sum last row
lastrow_sum <- sum(matrix1[n_dims,]) 
print(lastrow_sum)
 
```

Read about the eigen() function and use it on your matrix
```{r, eval=TRUE, echo=TRUE}
matrix_eigen <- eigen(matrix1) 
print(matrix_eigen)
```

Dig in with the typeof() function to figure out their type.
```{r, eval=TRUE, echo=TRUE}
typeof(matrix_eigen$values)
```

```{r, eval=TRUE, echo=TRUE}
typeof(matrix_eigen$vectors)
```

### 2
Create a list with the following named elements:

```{r, eval=TRUE, echo=TRUE}
# my_matrix, which is a 4 x 4 matrix filled with random uniform values

my_matrix <- matrix(data=runif(16), nrow=4)
print(my_matrix) 
```

```{r, eval=TRUE, echo=TRUE}
# my_logical which is a 100-element vector of TRUE or FALSE values. Do this efficiently by setting up a vector of random values and then applying an inequality to it.
my_logical <- (runif(100) > .8)
print(my_logical) 
```

```{r, eval=TRUE, echo=TRUE}
# my_letters, which is a 26-element vector of all the lower-case letters in random order.
my_letters <- sample(letters)
print(my_letters)
```

```{r, eval=TRUE, echo=TRUE}
# create a new list, which has the element[2,2] from the matrix, the second element of the logical vector, and the second element of the letters vector
my_list1 <- list(my_matrix, my_logical, my_letters)

new_list <- list(my_matrix[2,2], my_logical [2], my_letters[2]) 
print(new_list)
```

```{r, eval=TRUE, echo=TRUE}
# use the typeof() function to confirm the underlying data types of each component in this list
typeof(new_list[[1]])
```

```{r, eval=TRUE, echo=TRUE}
typeof(new_list[[2]])
```

```{r, eval=TRUE, echo=TRUE}
typeof(new_list[[3]])
```

```{r, eval=TRUE, echo=TRUE}
# combine the underlying elements from the new list into a single atomic vector with the c() function
combined_elements_vector <- c(new_list[[1]], new_list[[2]], new_list[[3]])
print(combined_elements_vector)
```

```{r, eval=TRUE, echo=TRUE}
# what is the data type of this vector?
typeof(combined_elements_vector)
```

### 3
Create a data frame with the two variables (= columns) and 26 cases (= rows) below:

```{r, eval=TRUE, echo=TRUE}
#call the first variable my_unis and fill it with 26 random uniform values from 0 to 10

# call the second variable my_letters and fill it with 26 capital letters in random order.
dataframe <- data.frame(my_unis = runif(26, min=0, max=10), my_letters = sample(LETTERS))
print(dataframe)
```

```{r, eval=TRUE, echo=TRUE}
# for the first variable, use a single line of code in R to select 4 random rows and replace the numerical values in those rows with NA.
dataframe[sample(1:26, size=4), 1] <- NA
```

```{r, eval=TRUE, echo=TRUE}
# for the first variable, write a single line of R code to identify which rows have the missing values
which(is.na(dataframe$my_unis))
```

```{r, eval=TRUE, echo=TRUE}
# re-order the entire data frame to arrange the second variable in alphabetical order
dataframe <- dataframe[order(dataframe$my_letters),]
print(dataframe)
```

```{r, eval=TRUE, echo=TRUE}
# calculate the column mean for the first variable
mean(dataframe$my_unis, na.rm=TRUE)
```

