---
title: "Homework7_NZ"
author: "NZ"
date: "2025-03-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This homework assignment focus on data manipulation in R. Complete these problems using the dplyr and tidyverse packages.

### 1. Examine the structure of the iris data set. How many observations and variables are in the data set?

```{r, eval=TRUE, echo=TRUE}
library(tidyverse)
```

```{r, eval=TRUE, echo=TRUE}
str(iris)
```

### 2. Create a new data frame iris1 that contains only the species virginica and versicolor with sepal lengths longer than 6 cm and sepal widths longer than 2.5 cm. How many observations and variables are in the data set?

```{r, eval=TRUE, echo=TRUE}
iris1 <- iris %>% filter(Species == c("versicolor", "virginica"), Sepal.Length > 6.0, Sepal.Width > 2.5)
str(iris1)
```

### 3. Now, create a iris2 data frame from iris1 that contains only the columns for Species, Sepal.Length, and Sepal.Width. How many observations and variables are in the data set?

```{r, eval=TRUE, echo=TRUE}
iris2 <- iris1 %>% select(Species, Sepal.Length, Sepal.Width)
str(iris2)
```

